A new tidy data structure to support exploration and modeling of temporal data

### 2019-08-29 tsibble.Rmd
re-compile

A closer examination of some big airports across the US will give an indication of how well the busiest airports manage the outflow traffic on a daily basis. A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seattle (SEA) airports is obtained first. The succeeding operations compute delayed percentages every day at each airport, which are shown as grey lines in Figure \@ref(fig:sel-monthly-plot). Winter months tend to fluctuate a lot compared to the summer across all the airports. Superimposed on the plot are two-month moving averages, so the temporal trend is more visible. Since the number of days for each month is variable, moving averages over two months will require a weights input. But the weights specification can be avoided using a pair of commonly used rectangling verbs--`nest()` and `unnest()`, to wrap data frames partitioned by months into list-columns. The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year, for IAH, JFK, LAX and SEA. LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady delay. The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year. This pipeline gets the data into the daily series, and shifts the focus to five major airports.

@@

@@

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time in the region, organized into a calendar display (Figure \@ref(fig:calendar-plot)) using the **sugrrants** package [@wang-calendarviz-2018]. Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes the dominant usage pattern. This is probably driven by the use of air conditioning, because high peaks mostly correspond to hot days, where daily average temperature greater than 25 degrees Celsius. In the winter time (July and August) the daily pattern sees two peaks, which is probably due to heating in the morning and evening.

@@

(ref:elec-forecast-cap) One-day (48 step ahead) forecasts generated by ARIMA models, with and without temperatures, plotted against the actual demand. Both nicely capture the temporal dynamics, but ARIMA with temperature performs better than the model without.

>>>>

@@

A closer examination of some big airports across the US will give an indication of how well the busiest airports manage the outflow traffic on a daily basis. A subset that contains observations for Houston (IAH), New York (JFK), Kalaoa (KOA), Los Angeles (LAX) and Seattle (SEA) airports is obtained first. The succeeding operations compute delayed percentages every day at each airport, which are shown as grey lines in Figure \@ref(fig:sel-monthly-plot). Winter months tend to fluctuate a lot compared to the summer across all the airports. Superimposed on the plot are two-month moving averages, so the temporal trend is more visible. Since the number of days for each month is variable, moving averages over two months will require a weights input. But the weights specification can be avoided using a pair of commonly used rectangling verbs--`nest()` and `unnest()`, to wrap data frames partitioned by months into list-columns. The sliding operation with a large window size smooths out the fluctuations and gives a stable trend around 25% over the year, for IAH, JFK, LAX and SEA. LAX airport has seen a gradual decline in delays over the year, whereas the SEA airport has a steady delay. The IAH and JFK airports have more delays in the middle of year, while the KOA has the inverse pattern with higher delay percentage in both ends of the year. This pipeline gets the data into the daily series, and shifts the focus to five selected airports.

@@

Aggregation across all individuals helps to sketch a big picture of the behavioral change over time in the region, organized into a calendar display (Figure \@ref(fig:calendar-plot)) using the **sugrrants** package [@wang-calendarviz-2018]. Each glyph represents the daily pattern of average residential electricity usage every thirty minutes. Higher consumption is indicated by higher values, and typically occurs in daylight hours. Color indicates hot days. The daily snapshots vary depending on the season in the year. During the summer months (December and January), the late-afternoon peak becomes the dominant usage pattern. This is probably driven by the use of air conditioning, because high peaks mostly correspond to hot days, where daily average temperatures are greater than 25 degrees Celsius. In the winter time (July and August) the daily pattern sees two peaks, which is probably due to heating in the morning and evening.

@@

(ref:elec-forecast-cap) One-day (48 steps ahead) forecasts generated by ARIMA models, with and without temperatures, plotted against the actual demand. Both nicely capture the temporal dynamics, but ARIMA with temperature performs better than the model without.
